---
title: "Assignment 1a (R) - refactor obesity dataset code"
output: html_notebook
---

The below code downloads data from the obesity dataset from the [UC Irvine ML repository](https://doi.org/10.24432/C5H31Z). It then preprocesses the data, fits a machine-learning model and shows results. The coding style intentionally is not perfect.

In this assignment you'll refactor this code using the principles taught in the Reproducible Coding practices course. Things to consider are for example:

* Modularity
* Documentation
* Tests
* Packaging
* Dependencies
* Version control
* etc.

For the full assessment criteria see the Rubric. For full details on the assignment see the `README.md`

```{r Set up Environment and Load PAckages, echo=F}
# Virtual Environment Setup for R
install.packages('renv', repos='http://cran.us.r-project.org')
library(renv)
renv::init()

#Load necessary packages
renv::install('httr', repos='http://cran.us.r-project.org')
renv::install('readr', repos='http://cran.us.r-project.org')
renv::install('ggplot2', repos='http://cran.us.r-project.org')
renv::install('randomForest', repos='http://cran.us.r-project.org')
renv::install("styler")
renv::install("lintr")

library(httr)
library(readr)
library(ggplot2)
library(randomForest)
library(styler)
library(lintr)
```

```{r Load the Data, echo=F}
#Now we load the data from the UCI repository. We will download the zip file, extract it, and read the CSV file.

url <- 'https://archive.ics.uci.edu/static/public/544/estimation+of+obesity+levels+based+on+eating+habits+and+physical+condition.zip'

#DOwnload data and safe in data/
tmp <- GET(url, write_disk('data/data.zip', overwrite = TRUE))
httr::stop_for_status(tmp)

#unzip
utils::unzip('data/data.zip', exdir = './data')

obesity_raw <- read_csv('data/ObesityDataSet_raw_and_data_sinthetic.csv', show_col_types = FALSE)
```

```{r}
#Split data into the collected data and synthetic part (from row 499)
df_web <- obesity_raw[1:498, ]
df_synthetic <- obesity_raw[499:nrow(obesity_raw), ]
View(web_df)
```

```{r}
#inlcude BMI to both data sets
df_web$BMI <- bmi(df_web$Weight, df_web$Height)

df_synthetic$BMI <- bmi(df_synthetic$Weight, df_synthetic$Height)

# Assign labels to the web data set based on BMI categories

df_web$label <- 

outcomes <- c()
for (bmi in df_web$BMI) {
  if (bmi < 18.5) {
    outcomes <- c(outcomes, '0_undwerweight')
  } else if (bmi < 25) {
    outcomes <- c(outcomes, '1_normal')
  } else if (bmi < 30) {
    outcomes <- c(outcomes, '2_overweight')
  } else {
    outcomes <- c(outcomes, '3_obese')
  }
}
df_web$label <- outcomes

#plot the BMI of df_web as a histogram in each category
ggplot(df_web, aes(x = df_web$BMI)) +
geom_bar(fill = "lightblue") +
theme_minimal() +
labs(x = "BMI category", y = "Count")
```

```{r}
df_web[10:20, ]
```

```{r}
print(colnames(df))

df_web$MTRANS[df_web$MTRANS == 'Walking'] <- 1
df_web$MTRANS[df_web$MTRANS == 'Bike'] <- 1
df_web$MTRANS[df_web$MTRANS == 'Public_Transportation'] <- 0
df_web$MTRANS[df_web$MTRANS == 'Automobile'] <- 0
df_web$MTRANS[df_web$MTRANS == 'Motorbike'] <- 0
df_web$MTRANS <- as.numeric(df_web$MTRANS)

df_web$Gender[df_web$Gender == 'Male'] <- 1
df_web$Gender[df_web$Gender == 'Female'] <- 0
df_web$Gender <- as.numeric(df_web$Gender)

Xcol_binary <- c(
  'family_history_with_overweight',
  'FAVC',
  'SMOKE',
  'SCC'
)

for (col in Xcol_binary) {
  df_web[df_web[[col]] == 'yes', col] <- '1'
  df_web[df_web[[col]] == 'no', col] <- '0'
  df_web[[col]] <- as.numeric(df_web[[col]])
}

Xcol_ordinal <- c('CAEC', 'CALC')

df_web$CAEC <- as.numeric(factor(df_web$CAEC, levels = c('no', 'Sometimes', 'Frequently', 'Always'))) - 1
df_web$CALC <- as.numeric(factor(df_web$CALC, levels = c('no', 'Sometimes', 'Frequently', 'Always'))) - 1
```

```{r}
X_columns <- c('Age', 'Gender', 'family_history_with_overweight',
               'FAVC', 'FCVC', 'NCP', 'CAEC', 'SMOKE', 'CH2O', 'SCC', 'FAF', 'TUE',
               'CALC', 'MTRANS')
```




## Regression problem

```{r}
classes <- sort(unique(df_web$label))
freq <- table(df_web$label)
class_weight <- length(df_web$label) / (length(classes) * as.numeric(freq))
names(class_weight) <- names(freq)

print(classes)
print(class_weight)
```

```{r}
# Note: result can deviate from the python code due to random seed!
set.seed(42)
idx <- seq_len(nrow(df_web))

idx_train <- sample(idx, size = floor(0.7 * length(idx)))
idx_test <- setdiff(idx, idx_train)

X_train <- df_web[idx_train, X_columns]
X_test <- df_web[idx_test, X_columns]
y_train <- df_web$BMI[idx_train]
y_test <- df_web$BMI[idx_test]

sample_weights <- class_weight[df_web$label[idx_train]]
```

```{r}
reg <- lm(BMI ~ ., data = cbind(X_train, BMI = y_train))
reg <- randomForest(x = X_train, y = y_train, ntree = 100, nodesize = 5, maxnodes = 20)
```

```{r}
preds <- predict(reg, X_test)

plot(preds, y_test, pch = 16, col = rgb(0, 0, 0, 0.5))
abline(a = 0, b = 1, col = 'red', lty = 3)
```

```{r}
plot(predict(reg, X_train), y_train, pch = 16, col = rgb(0, 0, 0, 0.5))
abline(a = 0, b = 1, col = 'red', lty = 3)
```

```{r}
sample_weights_test <- class_weight[df_web$label[idx_test]]

mae_baseline <- sum(abs(y_test - rep(mean(y_train), length(y_test))) * sample_weights_test) / sum(sample_weights_test)
mae_model <- mean(abs(y_test - preds) * sample_weights_test)

cat(sprintf('Mean absolute error (baseline): %.3f\n', mae_baseline))
cat(sprintf('Mean absolute error: %.3f\n', mae_model))
```