---
title: "Assignment 1a (R) - refactor obesity dataset code"
output: html_notebook
---

The below code downloads data from the obesity dataset from the [UC Irvine ML repository](https://doi.org/10.24432/C5H31Z). It then preprocesses the data, fits a machine-learning model and shows results. The coding style intentionally is not perfect.

In this assignment you'll refactor this code using the principles taught in the Reproducible Coding practices course. Things to consider are for example:

* Modularity
* Documentation
* Tests
* Packaging
* Dependencies
* Version control
* etc.

For the full assessment criteria see the Rubric. For full details on the assignment see the `README.md`

```{r Set up Environment and Load PAckages, echo=F}
# Virtual Environment Setup for R
#renv::init()
#renv::activate()

#Load necessary packages
library(httr)
library(readr)
library(ggplot2)
library(randomForest)
library(styler)
library(lintr)
library(testthat)
library(tidyverse)
library(dplyr)
library(devtools)
library(testthat)
library(roxygen2)
library(reproduciblecode)

#source functions from my function folder
source(file.path("R", "Rfunctions.R"))
```

# Data description

This data set is retrieved from the UCI website. We have `498` subjects that have been surveyed and the rest is synthetically generated. THe data is in tabular format and describes for each subject `Demographics`, `Physical Attributes`, and other variables measured like `Smoking`. In this file we will explore the data and build a regression/Random Forest model to predict `BMI`. 

```{r Load the Data, echo=F}
#Now we load the data from the UCI repository. We will download the zip file, extract it, and read the CSV file.

url <- 'https://archive.ics.uci.edu/static/public/544/estimation+of+obesity+levels+based+on+eating+habits+and+physical+condition.zip'

#DOwnload data and safe in data/
tmp <- GET(url, write_disk('data/data.zip', overwrite = TRUE))
httr::stop_for_status(tmp)

#unzip
utils::unzip('data/data.zip', exdir = './data')

obesity_raw <- read_csv('data/ObesityDataSet_raw_and_data_sinthetic.csv', show_col_types = FALSE)
```

```{r}
#Split data into the collected data and synthetic part (from row 499)
df_web <- obesity_raw[1:498, ]
df_synthetic <- obesity_raw[499:nrow(obesity_raw), ]
View(df_web)
```

```{r}
#Inlcude BMI to both data sets

df_web$BMI <- bmi(df_web$Weight, df_web$Height)
df_synthetic$BMI <- bmi(df_synthetic$Weight, df_synthetic$Height)

# Assign labels to the web data set based on BMI categories
df_web <- df_web %>%
  mutate(label = case_when(
    BMI < 18.5 ~ "0_underweight",
    BMI < 25 ~ "1_normal",
    BMI < 30 ~ "2_overweight",
    BMI >= 30 ~ "3_obese"
  ))

#plot the BMI of df_web as a histogram in each category
ggplot(df_web, aes(x = label)) +
  geom_bar(fill = "lightblue") +
  theme_minimal() +
  labs(x = "BMI category", y = "Count")
```

```{r}
# Data restructuring

#Add the varibale "active" to the data set, where active is 1 if the person commutes by walk or bike, and 0 for the rest.
df_web <- df_web %>%
  mutate(MTRANS = ifelse(MTRANS %in% c("Walking", "Bike"), 1, 0))

df_web$MTRANS <- binary(df_web$MTRANS, c("Walking", "Biking"))

#Convert categorical variables to numeric.
df_web$Gender <- binary(df_web$Gender, "Male")
df_web$family_history_with_overweight <- binary(df_web$family_history_with_overweight, "yes")
df_web$FAVC <- binary(df_web$FAVC, "yes")
df_web$SMOKE <- binary(df_web$SMOKE, "yes")
df_web$SCC <- binary(df_web$SCC, "yes")


#Convert ordinal variables into numeric ones, starting from 0..
df_web <- df_web %>%
  mutate(CAEC = as.numeric(factor(CAEC, levels = c('no', 'Sometimes', 'Frequently', 'Always'))) -1,
         CALC = as.numeric(factor(CALC, levels = c('no', 'Sometimes', 'Frequently', 'Always'))) -1)
```






## Predicting BMI using the Survey Data

In this part we will take a look at how to predict BMI in the surveyed individuals. First we will look into class imbalance between the different `BMI` categories. For the `Random Forest` we will have to split our data into train and test sets, and define our `X` (predictors) and `Y` (outcome). Then we will run the random forest and extract the predictors.

```{r}
# Investigate class imbalance of BMI categories
classes <- sort(unique(df_web$label))
freq <- table(df_web$label)
class_weight <- length(df_web$label) / (length(classes) * as.numeric(freq))
names(class_weight) <- names(freq)

print(class_weight)
```

```{r}
# Split data into train/test with 70/30 split for the variables Gender, Age, and BMI
set.seed(42)
idx <- seq_len(nrow(df_web))

idx_train <- sample(idx, size = floor(0.7 * length(idx)))
idx_test <- setdiff(idx, idx_train)

#predictors
predictors <- c('Age', 'Gender', 'family_history_with_overweight',
               'FAVC', 'FCVC', 'NCP', 'CAEC', 'SMOKE', 'CH2O', 'SCC', 'FAF', 'TUE',
               'CALC', 'MTRANS')

#predictor train and test data
X_train <- df_web[idx_train, predictors]
X_test <- df_web[idx_test, predictors]

#outcome train and test data
y_train <- df_web$BMI[idx_train]
y_test <- df_web$BMI[idx_test]

#define sample weight for training data
sample_weights <- class_weight[df_web$label[idx_train]]
```

```{r}
#Run random forest on the training data
rf <- randomForest(x = X_train, y = y_train, ntree = 100, nodesize = 5, maxnodes = 20)
```

```{r}
preds <- predict(rf, X_test)

plot(preds, y_test, pch = 16, col = rgb(0, 0, 0, 0.5))
abline(a = 0, b = 1, col = 'red', lty = 3)
```

```{r}
plot(predict(reg, X_train), y_train, pch = 16, col = rgb(0, 0, 0, 0.5))
abline(a = 0, b = 1, col = 'red', lty = 3)
```

```{r}
sample_weights_test <- class_weight[df_web$label[idx_test]]

mae_baseline <- sum(abs(y_test - rep(mean(y_train), length(y_test))) * sample_weights_test) / sum(sample_weights_test)
mae_model <- mean(abs(y_test - preds) * sample_weights_test)

cat(sprintf('Mean absolute error (baseline): %.3f\n', mae_baseline))
cat(sprintf('Mean absolute error: %.3f\n', mae_model))
```